---
title: "MAT 4376 Project Four"
author: "Catherine Samson and Hanan Ather"
date: "26/10/2019"
output: 
  html_document:
    toc: true
---

# Global Cities PBI

## Introduction

The global cities dataset lists cities accross the world and gives different types of information about them. This information includes things like continent, city population size, life expectancy, etc... Of this information, the seven most important variables in this dataset are the country, the city population size, the gross domestic product, life expectancy, the unemployment rate, the city area, and the annual population growth. 
  The country of the city is the most important geographic indicator for this data because when comparing data on a global scale, it is much more useful to compare at a higher level than individual city. Furthermore, The city population size and area are both important because they give information on how large a city is, how much space there is, and therefore how crowded it is as well. For example, if a certain city has a much higher population size than another, but the area is much smaller, it gives a good idea as to how the city must be organized to support such numbers.
  The life expectancy and infant mortality also both can say a lot about a city. For example, life expectancy can be loosely tied with quality of life. This is because a poor and unhealthy life will lead to a shorter life expectancy. Therefore, this value is a good indicator of this characteristic in a city. Infant mortality will also say a lot on this subject. Furthermore, high infant mortality rates are often associated for poor and unhealthy regions. Even if a city has a high GDP, a low life expectancy and/or infant mortality rate can say a lot about how the majority of the population is living. 
  The unemployment rate is also a good indicator to learn more about an area, much like life expectancy, unemployment rates can say a lot about the quality of life in the city and how it may be for the people who actually live there. Finally, the gross domestic product of a city is a commonly used indicator for how the city is doing in economic terms. It can be used to measure the success of a city in terms different from the quality of life.
  
## Basic Visualizations

The first step involves reading in the data, and bringing in the libraries which will be used as well.

```{r}
# Reading in global cities data
globalCitiesData <- read.csv("GlobalCitiesPBI.csv")

# Bringing in neccessary libraries
library(knitr)
library(ggplot2)
library(stringr)
library(DMwR)
library(plotly)
```
The following below visualizations show multivariate plots of the global cities dataset.
Life expectancy versus GDP is shown using a scatterplot. This is because the data for both axes comes in a continuous form which is not connected by time. In order to show their relationship, each point must be individually plotted. Furthermore, the values have been organized by colour using the continent of origin of each city. This helps provide further context to the plot and elaborates on the the nature of the data and why some points may be higher or lower than others. It can also show potential connections of the relationship between life expectancy and GDP to their geographical region. Different locations have different socioeconomic factors, and these can help further explain the relationship between the two primary variables shown.
```{r}
# Plotting GDP and life expectancy using ggplot2
ggplot(data = globalCitiesData, mapping = aes(x = GDP.Per.Capita..thousands....PPP.rates..per.resident., y =Life.Expectancy)) + geom_point(aes(colour = Continent)) + xlab("GDP per Capita in Thousands") + ylab("Life Expectancy")+ ggtitle("Life Expectancy versus GDP") + theme_bw()
```
The plot of life expectancy versus unemployment rate is built in the same was as the one above which uses GDP. A scatter plot is optimal to display the potential relationship. Since the points can take on any value in a certain range, visualizations of the categorical kinds such as bar plots would not convery the information in the same effective manner.
```{r}
# Plotting unemployment rate and life expectancy
ggplot(globalCitiesData, mapping = aes(x = Life.Expectancy)) + geom_point(aes(y=Unemployment.Rate,colour=Continent))+ xlab("Life Expectancy") + ylab("Unemployment Rate")+ ggtitle("Life Expectancy versus Unemployment Rate") + theme_bw()
```
When visualizing a categorical and a continuous variable, bar plots can reveal a lot of information. The below plot gives the city population sizes for all the cities and has ordered them from largest to smallest. Finally, the legend also displays the Continent of each city once again. This visualization allows one to see which regions have the most populous cities as well as the sizes of each city exactly. A scatter plot in this scenario would show very similar data, but would not as effectively communicate the size differences in population per city due to the fact that this would be showed simply with the height of the point rather than with the dimensions of a bar of data as shown below.
```{r}
# Plotting city population size and continent
ggplot(globalCitiesData, aes(x=reorder(Geography,-City.Population..millions.),y=City.Population..millions.)) + geom_bar(stat="identity", aes(fill=Continent),colour="black", width=0.9)+theme(axis.text.x  = element_text(size=6.9)) + xlab("City Population Size") + ylab("City Population Size")+ ggtitle("City Population Sizes in Each Country")+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + theme_bw() # (Chang, 2009), (zero323, 2015)
```
This next plot shows the city areas and population sizes per continent. This information is useful to split up because different regions tend to have distinct architectures and styles of living as well as differing amounts of physical space to expand in. This means that splitting the data per continent can help show the information in a context where it is easier to compare each city to those in geopgraphically similar situations.
```{r}
# Plotting city areas and city population sizes split by continent
ggplot(globalCitiesData, aes(x = City.Area..km2., y =City.Population..millions., group=Continent)) + geom_point(aes(colour = Continent)) + xlab("City Area in Kilometers Squared") + ylab("City Population in Millions")+ ggtitle("City Areas versus City Population Sizes Accross Continents") + theme_bw()+facet_grid(~Continent)+theme(axis.text.x = element_text(angle = 90, hjust = 1)) # (Chang,2009)
```

## Definite Visualizations

The following plot shows the quality of the air within a city versus the GDP per Capita per person. This data was displayed in the form of a bubble plot since each city has values for air quality and gdp which both can take on any variable rather than being of a categorical nature. Furthermore, the size of each point changes based on the size of the population in each city. This helps provide further information and context to each value. 
The data for air quality was transformed in order to further clarify the message of the graph. The air quality index works by listing air quality on a scale of 0 to 500, with 0 being good and 500 being hazardous (US EPA, 2017). The problem with this is that using these values, one who is not familiar with the air quality index could believe that an extremely polluted city like Dehli for example has good air quality since it is listed as 200 rather than understanding that this value is actually a bad sign. Therefore, a new column sorting the air quality is created to reduce confusion.

The plot is interactable, hover over points for more information and double click on elements in the legend to isolate and examine them.
```{r}
# Plotting GDP and air quality with bubble size as city population and colour as continent
p <- plot_ly(globalCitiesData, x = ~GDP.Per.Capita..thousands....PPP.rates..per.resident., y = ~Air.Quality., type = 'scatter', mode = 'markers',
        hoverinfo = 'text',
        color = ~Continent,
        marker = list(size = ~City.Population..millions., sizeref = 0.6, showlegend=F),
        text = ~paste('</br> City: ', Geography,
                      '</br> Country: ', Country,
                      '</br> Average Life Expectancy in Years: ',  Life.Expectancy))%>%
        layout(title = "Air Quality versus GDP per Capita in Global Cities",
           xaxis = list(title = 'GDP per Capita'),
           yaxis = list(title = 'Air Quality Index')) %>%
      layout(yaxis = list(
      ticktext = list("Good", "Moderate", "Potentially Unhealthy", "Unhealthy", "Very Unhealthy"), 
      tickvals = list(50, 100, 150, 200, 300),
      tickmode = "array"))  # (Plotly, 2019)
p
```

The below plot shows the infant mortality rate in each city versus the life expectancy. For similar reasons as the above plot, the following plot uses a bubble plot form to show the relationship between two variables which are not categorical. This allows for the relationship between each axis to naturally be revealed by the position of the points. Furthermore, the size of the points now corresponds to the GDP of each city displayed while the colours still display the continent. This allows for further information to be communicated through the plot without adding another axis or more points. 
The plot is again interactable. Hover over points for further information. Additionally, if you wish to isolate certain elements in the legend, simply double click on them.
```{r}
# Plotting life expectancy and infant mortality rate with bubble size as gdp and colour as #continent
p <- plot_ly(globalCitiesData, x = ~ Life.Expectancy, 
             y = ~Infant.Mortality..Deaths.per.1.000.Births., type = 'scatter', mode = 'markers',
        hoverinfo = 'text',
        marker = list(size = ~GDP.Per.Capita..thousands....PPP.rates..per.resident., sizeref = 3, showlegend = T),
        color = ~Continent,
        text = ~paste('</br> City: ', Geography,
                      '</br> Gross Domestic Product in Thousands: ', GDP.Per.Capita..thousands....PPP.rates..per.resident.,
                      '</br> Life Expectancy in Years: ',  Life.Expectancy,
                      '</br> Infant Mortality per 1000 Births: ', Infant.Mortality..Deaths.per.1.000.Births.)) %>%
        layout(title = "Infant Mortality versus Life Expectency in Global Cities",
           xaxis = list(title = 'Average Life Expectancy in Years'),
           yaxis = list(title = 'Infant Mortality per 1000 Births')) # (plotly, 2019)
    
p
```

The below plot shows a way to combine scatter and line plots to easily convey more information. The plot gives the life expectancies for males and females for each city in the form of points and then also shows the average life expectancy per city as well. This average is graphed in the form of a line. This allows one to see whether points fall above or below the average with ease and therefore exemplifies the relationship found between life expectancy and gender.
```{r}
# Plotting life expectancy of males versus females as well as the average per city 
p<-ggplot(globalCitiesData) + geom_line(aes(x=Life.Expectancy,y=Life.Expectancy))+geom_point(aes(x=Life.Expectancy,y = Life.Expectancy.in.Years..Male.,colour="Male"))+geom_point(aes(x=Life.Expectancy,y = Life.Expectancy.in.Years..Female.,colour="Female"))+theme(axis.text.y  = element_text(size=6.8)) + xlab("Life Expectancy of Males in Years") + ylab("Life Expectancy of Females in Years")+ ggtitle("Life Expectancy of Females versus Males") + theme_bw()+scale_colour_manual("", 
                      breaks = c("Male", "Female"),
                      values = c("darkorchid1", "cornflowerblue")) # (Chang, 2009), (Diggs, 2012)
p<-ggplotly(p)
p
```
In conclusion, the data reveals many interesting observations. For example, it can be seen from the above plots that 

## Works Cited

"R Figure Reference" Plotly, 2019. https://plot.ly/r/reference/. Accessed October 25th, 2019.

Chang, Jonathon. "Rotating and Spacing Axis Labels in ggplot2". Stackexchange, August 25th, 2009. https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2. Accessed October 24th, 2019.

Diggs, Brian. "Add Legend to ggplot2 Line Plot". Stackexchange, April 27th, 2012. https://stackoverflow.com/questions/10349206/add-legend-to-ggplot2-line-plot. Accessed October 24th, 2019.

United States Environmental Protection Agency. "Air Quality Index". airnow.gov, July 27th, 2017. https://airnow.gov/index.cfm?action=aqi_brochure.index. Accessed October 24th, 2019.

zero323. "geom_bar Define Border Color with Different Fill Colors". Stackexchange, June 8th, 2015. https://stackoverflow.com/questions/30709972/geom-bar-define-border-color-with-different-fill-colors. Accessed October 24th, 2019.


# 2016 Collisions Data

## Introduction
In order to to visualize the data, we must first bring in the relevant libraries and the dataset itself.
```{r}
library(plotly)
library(dplyr)
library(ggplot2)
library(tidyverse)
```
```{r}
data = read.csv(file="C:/RData/2016collisionsfinalCSV.csv", header=TRUE, sep=",")
```
```{r}
head(data)
```

First off we can plot some of the variables to see if anything stands out. After analyzing visualizations the variables which seem important are: Collision Location,Environment,Road Surface, Traffic Controls, and Impact type. We came to this conclusion by plotting many different variables and analyzing the visulizations to see if we could find any patterns. To further understand the relationship between these variables we untilized multi-variates plots.

## Visualizations

```{r}
coll_chart_data= data %>%
  group_by(Collision_Location)  %>%
  summarise(count = n())
  plot_ly(coll_chart_data,labels = ~Collision_Location, values = ~count) %>%
  add_pie(hole = 0.6) %>%
    layout(title = "Collision Location",
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)) # (R Stat co., 2019), (Plotly, 2019)
```
As we can see most of the accidents according to our data set occur at non-intersection locations.
```{r}
## get rid of this plot (dougnout plot replaces it)
ggplot(data,aes(x = Collision_Classification)) +
  geom_bar(colour="purple", fill= "yellow",
            alpha=0.5,
           linetype="dashed",
           size=0.5) +
  labs(x = "Collision Classification") +
  labs(y = "Number of Accidents")+
  labs(title = "Collision Classification and Number of Accidents" ) # (Noamross, 2012)
  
```
This plot displays the 3 categories of collision classification present in this dataset and their respective counts per category. As we can see, majority of the collisions which were recorded in this dataset are P.D only (property damage). 

```{r}
ggplot(data,aes(x = Location)) +
  geom_bar(colour="purple", fill= "yellow",
            alpha=0.5,
           linetype="dashed",
           size=0.5)+
  labs(x = "Location") +
  labs(y = "Number of Accidents")+
  labs(title = "Location and Number of Accidents" ) # (EnvironmentComputing, 2019), (R Stat co., 2019), (Noamross, 2012)
  
```
The above plot looks at the distribution of the collisions locations. As we can notice, certian locations have a higher count of accidents occuring. 

Next, in order to get a more indepth view of the data, we will make a filter to look at the top 10 locations where most collisions occur in Ottawa.
```{r}
sum_enviro= data %>%
  group_by(Environment)  %>%
  summarise(count=n()) 
p <- plot_ly(sum_enviro, labels = ~Environment, values = ~count, type = 'pie') %>%
   layout(title = "Environment",
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)) # (Plotly, 2019), (R Stat co., 2019)
p
```
The above pie chart displays the relationship between the environmet and the number of the accidents which occured in that environment. As we can see, in this dataset most the the crashes seem to have occured in a clear environment. 

```{r}
p = ggplot(data,aes(x = Road_Surface)) +
  geom_bar(colour="green", fill= "blue",
            alpha=0.5,
           linetype="dashed",
           size=0.5) # (Noamross, 2012)
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(x = "Road Surface") +
  labs(y = "Number of Accidents")+
  labs(title = "Road Surface and Number of Accidents" )

```
The histogram above shows that most of the accidents that were recorded in this dataset occured in dry conditions. 
```{r}
colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)') # (Columbia U, n.d.)
sum_Traffic_Control= data %>%
  group_by(Traffic_Control)  %>%
  summarise(count=n()) 

sum_Impact_type= data %>%
  group_by(Impact_type)  %>%
  summarise(count=n()) 

sum_Light= data %>%
  group_by(Light)  %>%
  summarise(count=n()) 


p <- plot_ly() %>%

  add_pie(data = sum_Traffic_Control, labels = ~Traffic_Control, values = ~count,
          name = "Cut", domain = list(x = c(0, 0.4), y = c(0.4, 1))) %>%
  add_pie(data = sum_Impact_type, labels = ~Impact_type, values = ~count,
          name = "Color", domain = list(x = c(0.6, 1), y = c(0.4, 1))) %>%
  add_pie(data = sum_Light, labels = ~Light, values = ~count,
          name = "Clarity", domain = list(x = c(0.25, 0.75), y = c(0, 0.6))) %>%
  layout(title = "Traffic Controls, Impact type, and Light",showlegend = F,
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)) # (R Stat co., 2019), (Plotly, 2019)
p

```
Here we can visualize three differnt variables from the data set; Traffic Controls, Impact type, and Light.
We see that most the accidents occured when there were no traffic controls. Furthermore, the majority of the accidents occured in daylight. Addtionally we notice that the majority of the impact types are rear end hits. 

## Visualizations Multi-Variate

```{r}
p=ggplot(data, aes(x=Traffic_Control,y= Impact_type)) 
p+geom_bin2d() + theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.background = element_rect(fill = 'khaki'),
          panel.grid.major = element_line(colour = "burlywood", size=1.5),
          panel.grid.minor = element_line(colour = "tomato", 
                                          size=.25, 
                                          linetype = "dashed"),
          panel.border = element_blank(),
          axis.line.x = element_line(colour = "darkorange", 
                                     size=1.5, 
                                     lineend = "butt"),
          axis.line.y = element_line(colour = "darkorange", 
                                     size=1.5)) # (Columbia U, n.d.), (R Stat co., 2019)
```
This multivariate visualization compares the traffic control variable with the impact type varibale. This visulization shows that stop signs and rear ends have the highest accident count.
```{r}
p=ggplot(data, aes(x=Environment,y= Road_Surface)) 
p+ geom_count()+  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
theme(panel.background = element_rect(fill = 'white'),
          panel.grid.major = element_line(colour = "snow", size=1.5),
          panel.grid.minor = element_line(colour = "tomato", 
                                          size=.25, 
                                          linetype = "dashed"),
          panel.border = element_blank(),
          axis.line.x = element_line(colour = "indianred4", 
                                     size=1.5, 
                                     lineend = "butt"),
          axis.line.y = element_line(colour = "indianred4", 
                                     size=1.5)) # (Columbia U, n.d.), (R Stat co., 2019)
```
In the visulization above we examine the relationship between environment and road surface. As we can see the most of the accidents which occured happened in dry conditions and in clear weather. 
```{r}
p=ggplot(data, aes(x=Light, y= Collision_Location )) 
p+ geom_bin2d()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.background = element_rect(fill = 'white'),
          panel.grid.major = element_line(colour = "snow4", size=1.5),
          panel.grid.minor = element_line(colour = "tomato", 
                                          size=.25, 
                                          linetype = "dashed"),
          panel.border = element_blank(),
          axis.line.x = element_line(colour = "indianred4", 
                                     size=1.5, 
                                     lineend = "butt"),
          axis.line.y = element_line(colour = "indianred4", 
                                     size=1.5)) # (Columbia U, n.d.), (Stagraph, 2018), (R Stat co., 2019)

```
This visualization look at the relationship between Light and collisions location. We can see from the above visulization the highest count of accidents occur in daylight and at non-intersections.

```{r}
p=ggplot(data, aes(x=Light, y= Impact_type )) 
p+ geom_bin2d()+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.background = element_rect(fill = 'lavender'),
          panel.grid.major = element_line(colour = "darksalmon", size=1.5),
          panel.grid.minor = element_line(colour = "tomato", 
                                          size=.25, 
                                          linetype = "dashed"),
          panel.border = element_blank(),
          axis.line.x = element_line(colour = "darkorange4", 
                                     size=1.5, 
                                     lineend = "butt"),
          axis.line.y = element_line(colour = "darkorange4", 
                                     size=1.5)) # (Columbia U, n.d.), (Stagraph, 2018), (Plotly, 2019)
```
This bi-variate plot examines the relationship between light conditions versus the type of impact and the count of how many accidents were recorded in the dataset given those variables. We can see that most of the accidents that occured in this data set were rear end hits during the daylight. We should note that we cannot make any casual claims with this plot.  
```{r}
q =ggplot(data, aes(x=Light, y= Collision_Location)) # (Noamross, 2012)
q + geom_raster(aes(fill=Environment),hjust = 0.5,vjust = 0.5,interpolate = FALSE)+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # (R Stat co., 2019)
``` 
In this plot, collision location and light in the x-y and it overlays enviroment conditions overtop that. It looks at the relationship between collision location and light verus the environment. 
```{r}
p= plot_ly(data, x = ~Impact_type, y = ~Collision_Location) # (Plotly, 2019)
pp =subplot(
  p %>% add_markers(alpha = 0.2),
  p %>% add_histogram2d()
) # (R Stat co, 2019)
pp
```
This plot combines two differnt plots into one plot. It helps enhance out understand about the relationship between the impact type and collision location.

In conclusion, we can see that numerous factors lead higher amounts of collisions, however, correlation does not equal causation. This means that driving in these scenario does not neccessarily mean that you will be at danger since there are possibly other factors at play. 


## Works Cited

“A Quick Introduction to Ggplot().” Noamross.Net, 2012, www.noamross.net/archives/2012-10-05-ggplot-introduction/. Accessed 6 Nov. 2019.

“Colors in R.” Colombia University, n.d. http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf. Accessed 6 Nov, 2019.

“How to Geom_bin2d.” Stagraph.Com, 2018, stagraph.com/HowTo/Plot/Geometries/Geom/bin2d. Accessed 6 Nov. 2019.

“Modern Analytic Apps for the Enterprise.” Plotly, 2019, plot.ly/. Accessed 6 Nov. 2019.

“Plotting with Ggplot: Colours and Symbols – Environmental Computing.” Environmentalcomputing.Net, 2019, environmentalcomputing.net/plotting-with-ggplot-colours-and-symbols/. Accessed 6 Nov. 2019.

“The Complete Ggplot2 Tutorial - Part2 | How To Customize Ggplot2 (Full R Code).” R-Statistics.Co, 2019, r-statistics.co/Complete-Ggplot2-Tutorial-Part2-Customizing-Theme-With-R-Code.html#5.%20Faceting:%20Draw%20multiple%20plots%20within%20one%20figure. Accessed 6 Nov. 2019.

“Top 50 Ggplot2 Visualizations - The Master List (With Full R Code).” R-Statistics.Co, 2019, r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Scatterplot. Accessed 6 Nov. 2019.

# HR Census Simple Data

## Introduction

The HR 2016 Cenus data set contains over 105 variables and 127 observations. These observations range from looking at small subsections of the population, to  provinces, and one the observation is the also the entire canada. Therefore, we must use ratios to construct graphics that compare more than 2 variables. The first graph shows an example of what happens when we do not use ratios. Therefore, based on this, the most important variables for visual analysis are: Long-term unemployed Rate, Proportion of High school, proportion of Lone-parent families, Average total income in 2015, Proportion of Post-secondary graduates aged 25 to 54. We picked these variables because they are all ratios, so sample points of bigger regions are 'normalized' into ratios and secondly, it makes intuitive sense that there must exist some relationships between these variables.

```{r}
data = read.csv(file="C:/RData/HR_2016_Census_simple.csv", header=TRUE, sep=",")

# Bringing in neccessary libraries
library(knitr)
library(ggplot2)
library(stringr)
library(plotly)
head(data)
```

## Visualizations

The first variables we will look at are highschool graduation and unemployment rates. Naturally, one would think there should be some sort of relationship between these two variables.

```{r}
a = ggplot(data, aes(x= HSG, y = UE))

  a+geom_smooth(method="loess") +
labs( 
       y="Long-term unemployed", 
       x="High school graduates aged 25 to 29", 
       title="Scatterplot Comparing High school graduate vs.Long-term unemployed ", 
       caption = "Source: HR_2016_Census_simple")+
    geom_rug() # (RStudio, 2015)

```
As we can notice from the graph above there seems to be  a linear relationship between the number of high graduates and Long-term unemployment. This relationship seems to be a postive relationship which seems counter intutive. One would expect the data show show that education and number of unemployed people to be negatively correlated. A potential explanaition for why this graph may be doing is that the  observations come from regions which are highly varied in size. Naturally, regions with larger populations will have more people unemployed and more people that will graduate highschool. Thus this might be one reason why we are seeing this trend in the graph. However, it is important to note that after further visual investigation we can see that the variables in this dataset that are not normalized cannot be visulzied accurately.

```{r}
a = ggplot(data, aes(x= HSG_RATE, y = UE_RATE))

a+geom_smooth(method="loess") +
labs( 
       y="Long-term unemployed Rate", 
       x="High school graduates aged 25 to 29, Proportion", 
       title="Comparing High school graduate vs.Long-term unemployed Rates ", 
       caption = "Source: HR_2016_Census_simple")+
    geom_rug() # (RStudio, 2015)
```
This graph compares the Long-term unemployed Rate and Proportion of High school graduates aged 25 to 29. From this graph we can observe that highschool graduation rate is negatively correlated with long term unemployment. This is a correction from the previous graph. Once  we normalzie the data with respect to size, we can see that the highschool graduation rate and Long-term unemployment have a negative relationship. It is vital to note that we are not confusing a relationship with causation. However, in this dataset we choose the x-axis to be highschool graduation rate, which logically seemed more likely to be the independent variable within the context of this plot.
```{r}
packageVersion('plotly')
```

```{r}
x <- list(
  title = "Lone-parent families, proportion of census families"
  )

y <- list(
  title = "High school graduates aged 25 to 29, Proportion"
)
plot_ly(
  data, x = ~LONE_RATE, y = ~HSG_RATE,
  # Hover text:
  color = ~LONE_RATE, size = ~LONE_RATE) %>%
  layout(xaxis = x, yaxis = y) # (Plotly, 2019), (R Stat co., 2019)
```
As we can see from the above visulaization, there seems to be a trend between the proportion of lone-parent families and High school graduates aged 25 to 29. There is a negative correlation between the two variables within this dataset. However, this does not imply causation. 
```{r}
plot_ly(data, x = ~HSG_RATE, y = ~LONE_RATE, z = ~UE_RATE, color = ~HSG_RATE, 
        colors = c('#BF382A', '#0C4B8E')) %>%
        add_markers() %>%
        layout(scene = list(xaxis = list(title = 'High school graduates aged 25 to 29'),
                     yaxis = list(title = 'proportion Lone-parent families'),
                     zaxis = list(title = 'Long-term unemployed Rate'))) # (Plotly, 2019)
```
This graph displays the relationship between the three varibles we have explored in the dataset this far. The x-axis is the High school graduates rate, the y-axis is the proportion Lone-parent families and the z-axis the long-term unemployment rate. The trend we can notice from this graphic is that Low rate of highscool graduation and high proportion Lone-parent families rate have high unemploymentent rate. However, we are not making the mistake of confusing the correlation with causation, there could be other factors which explain the the distribution of the data.
```{r}
plot_ly(data, x = ~HSG_RATE, y = ~AVE_PERS_INC,
        marker = list(size = 10,
                       color = 'rgba(255, 182, 193, .9)',
                       line = list(color = 'rgba(152, 0, 0, .8)',
                                   width = 2))) %>%
  layout(title = "High school graduates aged 25 to 29 vs. Average total income in 2015 ",
         xaxis = list(title = "High school graduates aged 25 to 29"),
         yaxis = list (title = "Average total income in 2015")) # (R Stat co., 2019)
```
In this graph  High school graduates aged 25 to 29 was chosen as the independent variable and we want to examine if there is some relationship between the Average total income in 2015 that is present in this dataset. We can see that the variance of the average income increases a lot when propotion of highschool gradudates is above 80%. This makes logical sense because some proportion highschool graudates will go on to university and others may not, and this may result in the variation of average total income. However, there may be even more variables to consider and we should not draw any strong conclusions. 
```{r}
a = ggplot(data, aes(x= PSG_RATE, y = UE_RATE))

a+geom_smooth(method="loess") +
labs( 
       y="Long-term unemployed Rate (%)", 
       x="Proportion of Post-secondary graduates aged 25 to 54 (%)", 
       title="Comparing Post-secondary graduates vs.Long-term Unemployement Rate", 
       caption = "Source: HR_2016_Census_simple")+
    geom_rug() # (RStudio, 2015)
```
This graph compares the post-secondary graduates rate versus the long-term unemployemment rate. Overall, we can see that these two variables do have a negative relationship. Although it does contain region at (x=60%) where as the proportion of post-secondary graduates increaese the un-employment rate does as well. However, we need to consider the fact that this data is not perfect; as the observations in this data vary widely, and secondly there are many addtional variables which could be influencing unemployement rate. However, overall this graph gives us the impression that proportion of post-secondary and unemployment rate have a negative, 'linear' relationship. 
```{r}
plot_ly(data, x = ~ PSG_RATE, 
        y = ~UE_RATE, 
        z = ~AVE_PERS_INC, 
        color = ~HSG_RATE, 
        colors = c('#BF382A', '#0C4B8E')) %>%
        add_markers() %>%
        layout(scene = list(xaxis = list(title = 'Proportion of Post-secondary graduates'),
                     yaxis = list(title = 'Long-term Unemployement Rate'),
                     zaxis = list(title = 'Average total income in 2015'))) # (Plotly, 2019)
```
This plot displays the Proportion of Post-secondary graduates on the x-axis, Long-term Unemployement Rate on the y-axis and the Average total income in 2015. As we can see observations with the lowest long-term unemployment and the highest proportion of post-secondary graduates have the highest average income.

## Works Cited

“3D Scatter Plots.” Plot.Ly, 2019, plot.ly/r/3d-scatter-plots/. Accessed 6 Nov. 2019.

“Plotly Open Source Graphing Libraries.” Plot.Ly, 2019, plot.ly/graphing-libraries/. Accessed 6 Nov. 2019.

“Scatter and Line Plots.” Plot.Ly, 2019, plot.ly/r/line-and-scatter/. Accessed 6 Nov. 2019.

RStudio. "Graphical Primitives Data Visualization with Ggplot2 Geoms -Use a Geom to Represent Data Points, Use the Geom’s Aesthetic Properties to Represent Variables. Each Function Returns a Layer. One Variable Geoms". RStudio, March, 2015. https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf. Accessed November 6th, 2019.

R Stat co. “Top 50 Ggplot2 Visualizations - The Master List (With Full R Code).” R-Statistics.Co, 2019, r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html. Accessed November 6h, 2019.